Programming Language Survey Paper on Haskell

### 2013-04-22 projects/term_paper/paper.tex
Standardized code references and made a few edits through 2.2.2.

The data type for each parameter of a function can be explicitly declared by using the ``::" symbol, which is 
read as ``has a type of." The types of the parameters can be listed in order, and the last type in the list 
is the return type. The following function is called ``greet."

It takes in a string and returns a string. The string is passed in as the variable ``name." Everything that follows the equal sign is the expression, and the evaluation of this expression is the return value.

@@

Bindings in Haskell can be created in two different ways. The first is using `where' bindings. Where bindings allow 
the definition of a function inside another function. The scope of the function declared by a where is locally bound 
by the function that created it. Functions outside of the declaring function cannot call the function declared inside 
the where. Additionally, where bindings can be nested; that is, a function declared inside a where can declare its 
own helper functions inside a where. Where bindings can also span across guards, which have not been discussed yet.

In the following example, the function rectPrismVol declares two functions: faceArea and baseArea. Only the rectPrismVol 
function is allowed to call these functions because they are locally bound.

@@

The second type of binding available in Haskell is done using `let' bindings. Let bindings allow the declaration of 
variables that are expressions themselves and are locally bound. Let bindings are often used for pattern matching. 
The syntax of let bindings involved declaring the bindings inside a `let' declaration, followed by defining an 
expression inside an `in' declaration. The let and in statements must be aligned in the same column.

@@

The difference between let and where bindings is that let bindings are expressions themselves, whereas where 
bindings are just syntactic constructs. Additionally, unlike where bindings, let bindings cannot span across guards, 
a feature that has not yet been discussed.

@@

All operators in Haskell take two arguments, and they are usually called using infix notation; that is, the operator is 
placed between its two arguments. Although it is possible to use operators in prefix notation, it is customary to use 
infix notation to increase readability. This contrasts from most functional programming languages.

@@

\begin{minted}{haskell}
(*) :: (Num a) => a -> a -> a
(*) a b = a * a * b
print (2 * 3)   -- this yields 12 rather than 6
\end{minted}

>>>>

The data type for each parameter of a function can be explicitly declared by using the \texttt{::} symbol, which is 
read as ``has a type of.''  The types of the parameters can be listed in order, and the last type in the list 
is the return type. The following function is called \texttt{greet}.

It takes in a string and returns a string. The string is passed in as the variable \texttt{name}. Everything that follows 
the equal sign is the expression, and the evaluation of this expression is the return value.

@@

Bindings in Haskell can be created in two different ways. The first is using \texttt{where} bindings. \texttt{where} bindings allow 
the definition of a function inside another function. The scope of the function declared by a \texttt{where} is locally bound 
by the function that created it. Functions outside of the declaring function cannot call the function declared inside 
the \texttt{where}. Additionally, \texttt{where} bindings can be nested; that is, a function declared inside a \texttt{where} can declare
its own helper functions inside a \texttt{where}. \texttt{where} bindings can also span across guards, which will be covered shortly.

In the following example, the function \texttt{rectPrismVol} declares two functions: \texttt{faceArea} and \texttt{baseArea}. Only the 
\texttt{rectPrismVol} function is allowed to call these functions because they are locally bound.

@@

The second type of binding available in Haskell is done using \texttt{let} bindings. \texttt{let} bindings allow the declaration of 
variables that are expressions themselves and are locally bound.  \texttt{let} bindings are often used for pattern matching. 
The syntax of \texttt{let} bindings involved declaring the bindings inside a \texttt{let} declaration, followed by defining an 
expression inside an \texttt{in} declaration. The \texttt{let} and \texttt{in} statements must be aligned in the same column.

@@

The difference between \texttt{let} and \texttt{where} bindings is that \texttt{let} bindings are expressions themselves, whereas 
\texttt{where} bindings are just syntactic constructs. Additionally, unlike \texttt{where} bindings, \texttt{let} bindings cannot 
span across guards.

@@

All operators in Haskell are binary, and they are usually called using infix notation; that is, the operator is 
placed between its two arguments. Although it is possible to use operators in prefix notation, it is customary to use 
infix notation to increase readability. This contrasts from most functional programming languages.

@@

\begin{minted}{haskell}
(*) :: (Num a) => a -> a -> a
(*) a b = a + a + a + a
print (2 * 3)   -- yields 8 rather than 6
\end{minted}
